That's No Moon: New Analysis Shows No Evidence for Lunar Companion Orbiting Kepler-1625b

### 2019-01-23 ms.tex
discussion

\begin{abstract}
    Observations of the Kepler-1625 system with the Kepler and Hubble Space Telescopes have suggested the presence of a candidate exomoon, a Neptune-radius satellite orbiting a long-period Jovian planet. Here we present a new analysis of the Hubble observations from an independent data reduction pipeline. We find that the transit light curve is well fit with a planet-only model, and that the addition of a moon does not significantly improve the fit quality. The $3\,\sigma$ upper limit on the moon radius is $4.0R_\oplus$. FIXME: why are we different from TK18?  We conclude that FIXME, and future follow-up efforts for exomoons should focus on alternative targets.
\end{abstract}

@@

To begin the data reduction, we fit the centroid of the direct image with a two-dimensional Gaussian. The centroid position determines the position of the spectral trace, which we calculated using the coeffients provided in the configuration file from STScI: \texttt{G141.F130N.V4.32.conf}\footnote{available at \url{http://www.stsci.edu/hst/wfc3/analysis/grism_obs/calibrations/wfc3_g141.html}}.  To process the spectra, we flatfielded the raw data using the spectroscopic flatfield coefficients provided by STScI in \texttt{WFC3.IR.G141.flat.2.fits}, following the instructions in Section 6 of the aXe User Manual\footnote{\url{http://axe-info.stsci.edu/}}.  We then created an extraction box centered on the spectral trace. We varied the height and width of the box in 1-pixel increments to find the window that minimized the root-mean-square (rms) deviation from the best fit to the transit light curve.  The best was $445  < \mathrm{X} < 569$, and $517 < \mathrm{Y} < 531$, where X and Y are physical pixels in the spectral and spatial direction, respectively.

We reduced the grism exposures with the optimal extraction routine of \cite{horne86}, which minimizes background noise in the extracted spectrum by weighting pixels that are dominated by the target spectrum more heavily than pixels dominated by the background.  The inputs for optimal extraction are the background-subtracted data array, the error array (including photon noise read noise, and uncertainty due to background subtraction), an initial guess for the spectrum and its uncertainty, and a mask array for bad pixels.  For the initial guess of the spectrum and its uncertainty, we did a simple box extraction (sum over all rows in the extraction window), and assumed the variance was equal to the box-extracted spectrum (expected for photon noise limit).  We measured and subtracted the background from the data array as described in \ref{sec:background}. For the error array, we used a quadrature sum of the photon noise (the square root of the pixel counts), the read noise (12 photoelectrons for \texttt{flt} files; WFC3 Data Handbook\footnote{\url{http://www.stsci.edu/hst/wfc3/documents/handbooks/currentDHB/}}), and the error due to background subtraction (described in \ref{sec:background}).  The initial bad pixel mask included pixels marked with the data quality flag 4 or 512 (dead pixels or blobs).

In brief, optimal extraction is an iterative procedure with the following steps. First, we created a smoothed image by median-filtering each row of the data with a 9-pixel-wide window.  We then normalized the smoothed image by dividing each column by its sum, and multiplied it by the best guess spectrum. We compared the smoothed image to the real data and masked outliers in the data that are greater than a threshold $\sigma_\mathrm{cut}$. We then recomputed the best guess spectrum with the new mask and the optimal weights from \cite{horne86}. The process is iterated until no outliers greater than the threshold remain.  This procedure masks any cosmic rays or bad pixels that were missed by the initial \texttt{flt} calibration.  We tested a range of thresholds ($4 < \sigma_\mathrm{cut} < 15$) and found that different $\sigma_\mathrm{cut}$ choices did not significantly change the final transit light curve. The data reported in this work use $\sigma_\mathrm{cut} = 15$.  To create the broadband transit light curve, we sum each spectrum over all wavelengths.

@@

\subsection{Background Subtraction} 
\label{sec:background} 
The star Kepler-1625 is faint (H mag = 14.0) relative to most other exoplanet host stars observed with WFC3, which makes accurate background subtraction especially important for this target. Moreover, the host star is in a crowded field, so the pixels used to estimate the background must be chosen carefully to avoid contamination from other stars.  To estimate the background counts, we masked pixels with total counts larger than 800 electrons and took the median count in the unmasked pixels. The per pixel uncertainty due to background subtraction is 1.4826 times the median absolute deviation.

@@

\section{Results}
We obtained an excellent fit to the light curve with the no-moon model, as illustrated in  Figure\,\ref{fig:bestfit}. The residuals to our no-moon model fit have rms equal to 362.2 parts per million (ppm), which is within 2\% of the predicted photon shot noise (367 ppm), and yields a $\chi_\nu^2 = 1.05$. The binned rms decreases with the square root of the number of points per bin, as expected for photon noise-limited statistics (see rms versus bin size in Figure~\ref{fig:rms}).

\begin{figure}
\includegraphics[width = 0.5 \textwidth]{figures/fig2_rms.pdf}
    \caption{Light curve rms versus bin size for the best fit no-moon model. The fit to data from this work (blue line) agrees well with the expected photon-limited, $\sqrt{N}$ decrease in rms with bin size (black line). The TK18 rms (red line) ranges from $1.3 - 2\times$ the photon limit for bin sizes of 1 to 20 data points. The increase in relative rms for larger bin sizes indicates that time-correlated noise is present in the TK18 data.}
\label{fig:rms}
\end{figure}

We also obtained a good fit with the moon model; however, the addition of three more free parameters does not significantly improve the fit quality. The best fit moon model has a radius $r_\mathrm{moon} = 2.4R_\oplus$, a mid-transit time $t_\mathrm{moon} = 2458055.2869\,\mathrm{BJD_{TDB}}$, and a period $P_m = 26.5$ days. This fit has a slightly lower rms than the no-moon model (359.5 ppm versus 362.2 ppm), but the $\chi_\nu^2$ is unchanged (also $1.05$). According to the Bayesian information criterion (BIC), which penalizes unnecessary model complexity, the moon model is disfavored with $\Delta\mathrm{BIC} = 13$. This constitutes strong evidence against the inclusion of a moon \citep{kass95}.  In addition, the moon transit time is not well-constrained by the data. As shown in the posterior distribution in Figure~\ref{fig:corner}, the $2\,\sigma$ confidence interval for the moon's transit includes the entire visit.  The upper limit on the moon radius is $3.2\,R_\oplus$ at $2\sigma$ confidence and $4.0\,R_\oplus$ at $3\sigma$.

@@

We fit the data with both the no-moon and models.  For the systematics, we tested the XY decorrelation model described above, but found that it did not perform as well as the systematics models presented in TK18 (10\% higher rms and time-correlated residuals).  We therefore opted to use a TK18 systematics model to enable a fair comparison between our results. We used a second-order polynomial trend in time, a constant offset after orbit 14 (where the guide stars are reacquired), and the non-parameteric orbit-long ramp model. This model has one more free parameter than the systematics model given in Equation~\ref{eq:sys}. Figure\,\ref{fig:bestfit} shows the best fit models.

Overall, we find that the TK18 fits have a larger rms than the fits to our new data reduction. The no-moon model has an rms of 381.5 ppm (compared to 362.2 for our data). As TK18 found, the moon model improves the rms to 363.4 ppm ($\Delta\chi^2 = 22.0$); however, even with the additional four free parameters, the moon model does not perform as well as the no-moon model fit to our data.  This difference in fit quality corresponds to $\Delta\mathrm{BIC} = 23.9$, which is strong evidence against the moon.  
%Even when a moon is included, the fit is still slightly worse (rms of 363.4 ppm) than our best no-moon model, in spite of four additional free parameters.

The moon model also yields qualitatively different posterior distributions for the two data sets.  For the TK18 data, the moon radius and transit time are peaked at $r_\mathrm{moon} = 4.0^{+0.8}_{-0.6}\,R_\oplus$ ($>2.1\,R_\oplus$ at $3\sigma$ confidence), and $t_\mathrm{moon} =  2458056.29^{+0.06}_{-0.04}\,\mathrm{BJD_{TDB}}$ (as shown in Figure\,\ref{fig:corner}). By contrast, the fit to the new data only yields an upper limit on the moon radius of ($4.0\,R_\oplus$ at $3\sigma$ confidence), and the transit time is unconstrained.

%Our results are consistent with those originally published in TK18, including the large improvement in $\chi^2$ when the moon is included ($\Delta\chi^2 = 22.0$). The posterior distribution for

%We obtain a better fit (a lower $\chi^2$) with the no-moon model fit to our data than we obtain with the moon model fit to the TK18 data, even though it has four additional free parameters.

\begin{figure}
\includegraphics[width = 0.5 \textwidth]{figures/fig5_pairs.pdf}
    \caption{Posterior distributions for the moon radius and time of transit based on an MCMC fit to data from this work (blue) and from TK18 (red). The shading corresponds to $1-$, $2-$, and $3\sigma$ contours (darkest to lightest). These values are marginalized over all other model parameters. The diagonal shows the normalized posterior probability density.}
\label{fig:corner}
\end{figure}

\section{Discussion}
As discussed above, we do not recover the moon. Our analysis was designed to be independent, and as such makes different choices about the data reduction and the choice of sytematics model.

We explored several avenues to determine what caused the difference in our results. We first checked whether the detrending model was responsible.

We also compared the data directly (see Figure\,\ref{fig:raw}). The data is different!
rotate the image, aperture choice, measurement of the background, outlier masking,

In this work, we presented an independent analysis of HST observations of the Kepler-1625 system. In contrast to previous analysis by \citep{teachey18b}, we do not find evidence for the transit of a moon. We obtain a better fit to the data without including a moon in the model.  We

>>>>

\begin{abstract}
    Observations of the Kepler-1625 system with the Kepler and Hubble Space Telescopes have suggested the presence of a candidate exomoon, Kepler-1625b-i, a Neptune-radius satellite orbiting a long-period Jovian planet. Here we present a new analysis of the Hubble observations from an independent data reduction pipeline. We find that the transit light curve is well fit with a planet-only model, with a best-fit $\chi^2_\nu$ equal to $1.01$. The addition of a moon does not significantly improve the fit quality. We compare our results directly with \cite{teachey18b}, and find that we obtain a better fit to the light curve using a model with fewer free parameters (no moon). We discuss possible sources for the discrepancy in our results, and conclude that there is no single explanation for the difference but rather that it is the sum of many small choices. We conclude that future follow-up efforts for exomoons should focus on alternate targets.

\end{abstract}

@@

To begin the data reduction, we fit the centroid of the direct image with a two-dimensional Gaussian. The centroid position determines the position of the spectral trace, which we calculated using the coeffients provided in the configuration file from STScI: \texttt{G141.F130N.V4.32.conf}\footnote{available at \url{http://www.stsci.edu/hst/wfc3/analysis/grism_obs/calibrations/wfc3_g141.html}}.  To process the spectra, we flatfielded the raw data using the spectroscopic flatfield coefficients provided by STScI in \texttt{WFC3.IR.G141.flat.2.fits}, following the instructions in Section 6 of the aXe User Manual\footnote{\url{http://axe-info.stsci.edu/}}.  We then created an extraction box centered on the spectral trace. We varied the height and width of the box in 1-pixel increments to find the window that minimized the root-mean-square (rms) deviation from the best fit to the transit light curve.  The best was $450  < \mathrm{X} < 574$, and $522 < \mathrm{Y} < 536$, where X and Y are physical pixels in the spectral and spatial direction, respectively.

We reduced the grism exposures with the optimal extraction routine of \cite{horne86}, which minimizes background noise in the extracted spectrum by weighting pixels that are dominated by the target spectrum more heavily than pixels dominated by the background.  The inputs for optimal extraction are the background-subtracted data array, the error array (including photon noise read noise, and uncertainty due to background subtraction), an initial guess for the spectrum and its uncertainty, and a mask array for bad pixels.  For the initial guess of the spectrum and its uncertainty, we did a simple box extraction (sum over all rows in the extraction window), and assumed the variance was equal to the box-extracted spectrum (expected for photon noise limit).  We measured and subtracted the background from the data array as described in \ref{sec:background}. For the error array, we used a quadrature sum of the photon noise (the square root of the pixel counts), the read noise (12 photoelectrons for \texttt{flt} files; WFC3 Data Handbook\footnote{\url{http://www.stsci.edu/hst/wfc3/documents/handbooks/currentDHB/}}), and the error due to background subtraction (described in \ref{sec:background}).  The initial pixel mask marked all pixels as good.

In brief, optimal extraction is an iterative procedure with the following steps. First, we created a smoothed image by median-filtering each row of the data with a 9-pixel-wide window.  We then normalized the smoothed image by dividing each column by its sum, and multiplied it by the best guess spectrum. We compared the smoothed image to the real data and masked outliers in the data that are greater than a threshold $\sigma_\mathrm{cut}$. We then recomputed the best guess spectrum with the new mask and the optimal weights from \cite{horne86}. The process is iterated until no outliers greater than the threshold remain.  This procedure masks any cosmic rays or bad pixels that were missed by the initial \texttt{flt} calibration.  We tested a range of thresholds ($7.5 < \sigma_\mathrm{cut} < 15$) and found that different $\sigma_\mathrm{cut}$ choices did not significantly change the final transit light curve. The data reported in this work use $\sigma_\mathrm{cut} = 7.5$.  To create the broadband transit light curve, we sum each spectrum over all wavelengths.

@@

\subsection{Background Subtraction} 
\label{sec:background} 
The star Kepler-1625 is faint (H mag = 14.0) relative to most other exoplanet host stars observed with WFC3, which makes accurate background subtraction especially important for this target. Moreover, the host star is in a crowded field, so the pixels used to estimate the background must be chosen carefully to avoid contamination from other stars.  To estimate the background counts, we masked pixels with total counts larger than 800 electrons (2.7 electrons/sec) and took the median count in the unmasked pixels. The per pixel uncertainty due to background subtraction is 1.4826 times the median absolute deviation.

@@

\section{Results}
We obtained an excellent fit to the light curve with the no-moon model, as illustrated in  Figure\,\ref{fig:bestfit}. The residuals to the no-moon model fit have rms equal to 356 parts per million (ppm), which is within 3\% of the predicted photon shot noise (367 ppm), and yields a $\chi_\nu^2 = 1.01$. The binned rms decreases with the square root of the number of points per bin, as expected for photon noise-limited statistics (see rms versus bin size in Figure~\ref{fig:rms}).

\begin{figure}
\includegraphics[width = 0.5 \textwidth]{figures/fig2_rms.pdf}
    \caption{Light curve rms versus bin size for the best fit no-moon model (solid lines) and moon model (dashed lines), for data from this work (red) and TK18 (blue). The fits to data from this work agree well with the expected photon-limited, $\sqrt{N}$ decrease in rms with bin size (black line).  We also reach the photon limit for the TK18 data, but only for the moon model. The rms for the no-moon fit (red dashed line) ranges from $1.1 - 1.5\times$ the photon limit for 1 to 20 points per bin.}
 \label{fig:rms}
\end{figure}

We obtained a slightly better fit with the moon model (rms = 351 ppm; $\chi_\nu^2 = 0.998$); however, this is not a large enough improvement in fit quality to merit the addition of three additional free parameters.  According to the Bayesian information criterion (BIC), which penalizes unnecessary model complexity, the moon model is disfavored with $\Delta\mathrm{BIC} = 10.7$. This constitutes strong evidence against the inclusion of a moon \citep{kass95}.  In addition, the moon transit time is not well-constrained by the data. As shown in the posterior distribution in Figure~\ref{fig:corner}, the $2\,\sigma$ confidence interval for the moon's transit includes the entire duration of the observations.  The upper limit on the moon radius is $3.2\,R_\oplus$ at $2\sigma$ confidence and $4.0\,R_\oplus$ at $3\sigma$.

%The best fit moon model has a radius $r_\mathrm{moon} = 2.4R_\oplus$, a mid-transit time $t_\mathrm{moon} = 2458055.2869\,\mathrm{BJD_{TDB}}$, and a period $P_m = 26.5$ days. This fit has a slightly lower rms than the no-moon model (359.5 ppm versus 362.2 ppm), but

@@

We fit the data with both the no-moon and models.  For the systematics, we tested the XY decorrelation model described above, but found that it did not perform as well as the systematics models presented in TK18 (10\% higher rms and time-correlated residuals).  We therefore opted to use a TK18 systematics model to enable a fair comparison between our results. We obtained the best fit using a second-order polynomial trend in time, a constant offset after orbit 14 (where the guide stars are reacquired), and the non-parameteric orbit-long ramp model. This model has one more free parameter than the systematics model given in Equation~\ref{eq:sys}. Figure\,\ref{fig:bestfit} shows the best fit models.

Similar to the findings of TK18, the moon model improved the fit quality by a $\Delta\chi^2 = 22.0$. Notably, however, the moon model fit to the TK18 data does not perform better than the no-moon model fit to our new data (rms of 363 versus 356 ppm), even with the additional four free parameters.  This difference in fit quality corresponds to $\Delta\mathrm{BIC} = 31.2$, which is strong evidence against the moon.

The moon model also yields qualitatively different posterior distributions for the two data sets.  As shown in Figure\,\ref{fig:corner}, for the TK18 data the moon radius and transit time are peaked at $r_\mathrm{moon} = 4.0^{+0.8}_{-0.6}\,R_\oplus$  and $t_\mathrm{moon} =  2458056.29^{+0.06}_{-0.04}\,\mathrm{BJD_{TDB}}$. By contrast, the fit to the new data presented here yields an upper limit on the moon radius of ($4.0\,R_\oplus$ at $3\sigma$ confidence), and the transit time is unconstrained.

\begin{figure}
\includegraphics[width = 0.5 \textwidth]{figures/fig5_pairs.pdf}
    \caption{Posterior distributions for the moon radius and time of transit based on an MCMC fit to data from this work (blue) and from TK18 (red). The shading corresponds to $1-$, $2-$, and $3\,\sigma$ contours (from darkest to lightest). These values are marginalized over all other model parameters.} 
\label{fig:corner}
\end{figure}

\section{Discussion}
A natural question arising from our results is what the difference is between TK18 and the new fits presented here. One possibility is that the source of the discrepancy is the raw data itself. As shown in Figure\,\ref{fig:raw}, our measured count rate ranges from $2.46 -2.74\%$ lower than the TK18 measurement, and there is a small bump in the difference between the two data sets near the location of the moon transit identified in the TK18 data (see the bottom panel).

 We explored several modifications to our pipeline to attempt to reproduce the TK18 reduction. These included rotating the image by 0.5 degrees, using the same aperture as TK18 to extract the spectrum, and scaling the master sky flat for the background subtraction (rather than just subtracting the median). None of these modifications had a significant effect on our results.

There are a few other steps in the TK18 data reduction that would require substantial modification of our pipeline to recreate, but seem unlikely to be responsible for the difference. One of these is outlier masking. TK18 identify outliers with a Gaussian process fit to the pixel-level light curves, compared to our optimal extraction approach. Despite the difference, both methods flag $\sim$0.01\% of pixels as bad.  We also do not use the STScI software \texttt{axeprep} to embed the raw $256\times256$ image in a larger array; however, this process primarily affects the edge of the image, many pixels distant from the extraction box, so it is unclear how this step would bias the light curve.

 %The only other difference we identify in the data reduction is the outlier masking. In our analysis, we iteratively masked outliers with the optimal extraction routine, whereas TK18 identified outliers via a Gaussian process fit to the pixel-level light curves. We flagged 0.07\% of pixels as bad, whereas TK18 flag twice as many. It is possible that this discrepancy could play role in the difference between the raw data sets.

In addition to these differences in data reduction, we also used a new detrending model. We opted for a decorrelation against X and Y position, motivated by known variation in the sensitivity of individual pixels on the detector and the presence of small shifts in the position of the spectrum (see Figure\,\ref{fig:shifts}). We applied this detrending model to the TK18 data, and obtained a poor fit to the light curve compared to the polynomial detrending (for the moon model with X-Y decorrelation, the rms was 404 ppm). We did not recover a large moon in this fit (the best fit was $r_\mathrm{moon} = 1.2R_\oplus$), but there was correlated noise present in the data that precluded a robust measurement of the uncertainty on the moon radius for this model. The poor quality of the fit points toward the raw data as the culprit for the difference in our results, but as we discuss, we could not identify a specific choice in the data reduction that is responsible.

 We conclude that no single choice in the data reduction or the modeling provides an easy explanation for the difference in our results, but rather that the difference is in the sum of many subtle choices.

\section{Conclusion}
The hunt for exomoons remains exciting and volatile. Our analysis of the exomoon candidate Kepler-1625b-i shows that the evidence in favor of the moon is highly sensitive to details of the data reduction and modeling. Future allocation of follow-up resources should prioritize alternative targets.

%\vspace{2cm}
